From Nested Virtualization Architecture \\To 2D Security Management for Distributed Clouds

### 2014-03-28 paper/draft.tex
updated section on architecture comparison.

\noindent The first analyzed architecture concerns only GPH. At first, it is important to recall the most interesting achievable by such architectures. In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers (horizontal requirements DP1).
Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH satisfies all horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.

To sum up, it could be possible to achieve live-migration between different platform and other features related to VM management flexibility but it could not still be possible to reduce the TCB. Analyzing MH approaches, the aim is try at most to retain to features, trying to introduce isolation and privacy improvements due to the lower TCB promised.

@@

\noindent MH on MH in reported only to see the orthogonality between this approach and GPH over GPH.
A theoretical implementation of MH over MH will provide the vertical requirements that are missing in the previous approach. The tiny core could be able to overcome monolithic problems due to privilege level of administration tasks (DP1), the possibility to improve the resilience of the platform to a large set of software failures (DP2) and finally reduce the possibility of the vulnerabilities of the most privilege part of the software (DP3). However, the interest on such architecture is limited because it is not able to overcome any cited issues of single-layer MH. Moreover, several horizontal properties are theoretically achievable but they will require a huge effort in developing functions of interest. Moreover, the lack of nested VMX functionalities prevents to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create a nested HVM guests.
To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not provide any improvement in this sense.

@@

\noindent The third architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
%Several MH are proposed in literature focusing on different .
%Therefore, this model tries to exploit microhypervisor as L1 hypervisor in order to enforce better isolation between L2 and L0. L1 will exploit the emulated nested VMX provided by GPH L0 and virtualize L2. 
%With this approach, we try to isolate the attacker in a VM virtualized on a secured layer, in order to prevent attacks to L0. The attacker would have to face with an essential virtualization layer and it would not be able to attack the back-end virtualization layer. 
An interesting point of this approach is the fact of being easily be deployed, without requiring any extra code addition to the GPH, L0.  Existing GPH-based platforms could potentially exports nested VMX features to those guests. Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the experimental status of those features.

However, the amount of code running in L0 kernel space is still as big as before and, therefore, no TCB reduction. 
Finally, adopting a MH as L1 prevents to use inter-platform enhanced features. As analyzed when dealing with MH design, the price of having a tiny core is normally paid with a lack of certain functionalities. For example, XMHF only supports one single guest. It could be easily noticed that this approach is not architecturally able to provide any new "horizontal" features, leveraging on nested virtualization. Therefore, the architecture is only able to provide the same features that are provided by today deployed ones, with an enhanced but not sufficient level of isolation.

@@

In this configuration, the TCB is smaller than other architectures, due to the modular structure of micro-kernels., executed at L0. Leveraging as L1 a GPH, like Xen or KVM, provides to the user an environment apparently  identical to the one used in production today, satisfying DP5. The L1 GPH is normally integrated with modern cloud toolkits like OpenStack and will support 3rd part applications.
Therefore, the architecture presents a stronger isolation leveraging on nested virtualization, providing in particular a strong tiny core(DP3). The MH layer will be transparent from the user and it will not not any change in the functioning of the platform.

@@

\section{Vertical Properties}

\section{Implementation}
In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction. In this section, we propose a GPH over MH implementation, based on Xen over NOVA \cite{nova}.

>>>>

\noindent The first architecture only features GPHs in both layers to serve as point of comparison. This type of architecture enables several remarkable features related to VM management flexibility such as live migration over different platforms, and transparent migration of virtualized clusters across data centers, fulfilling principle DP4. More generally, assuming complete and correct support of virtualization primitives for nested hardware in mainstream hypervisors, this architecture satisfies all horizontal design principles.

However, it fails to safisfy vertical design principles as TCB size is that of currently deployed NV architectures. The interest of other MH-based architectures will thus be to introduce isolation and privacy improvements by reducing the TCB, while still keeping horizontal features.

@@

\noindent This architecture is considered only to highlight improvements brought by a MH design compared to a GPH over GPH architecture. An MH over MH implementation would meet the vertical requirements missing in the previous approach. 
As in single-layer MH architectures, hypervisor disaggregation and minimization overcomes issues posed by monolithic system architectures and over-privileged administrative VMs (DP1). The smaller resulting TCB also improves resilience of the platform against a large set of software failures (DP2) and reduces the surface of attack of the most privileged part of the IaaS infrastructure (DP3).

However, the interest this class of architecture remains limited as it cannot overcome any of the generally cited issues of single-layer MHs such as compatibility with legacy hypervisor architectures (DP5). While several horizontal properties are theoretically achievable, they would require a large development effort to implement the relevant features such as interoperable advanced resource management. Moreover, the lack of nested VMX support prevents testing in practice this type of architecture, as MH usually rely on VMX primitives to create nested HVM guests.

Overall, with this architecture, if vertical challenges are well addressed as in single-layer MHs, it does not provide any improvement for horizontal challenges, despite NV being used to bridge gaps between IaaS platform eco-systems.

@@

\noindent The third architecture aims to alleviate security concerns. It combines a GPH as L0 and a MH as L1. 
%Several MH are proposed in literature focusing on different .
%Therefore, this model tries to exploit microhypervisor as L1 hypervisor in order to enforce better isolation between L2 and L0. L1 will exploit the emulated nested VMX provided by GPH L0 and virtualize L2. 
%With this approach, we try to isolate the attacker in a VM virtualized on a secured layer, in order to prevent attacks to L0. The attacker would have to face with an essential virtualization layer and it would not be able to attack the back-end virtualization layer. 
This approach is interesting as it may be easily deployed without requiring any extra code addition to the GPH.  Hardware support is not a real barrier as existing GPH-based platforms could potentially export nested VMX features to guests. Even if this is not yet the case in modern cloud platforms, this lack may be considered temporary due to the experimental status of such features.

However, there is no TCB reduction as the code size running in L0 kernel space is as big as before. 
Besides, adopting a MH as L1 prevents using inter-platform enhanced features. As already seen for MH design, the price of having a thin hypervisor is normally paid with a lack of certain features. For example, XMHF only supports a single guest~\cite{xmhf}. Thus, this architectural approach cannot provide any new horizontal features leveraging NV. Therefore, this approach can only provide the same features as those deployed today, with an increased but not insufficient level of isolation.

@@

\noindent In this configuration, the TCB is smaller than for other architectures, due to the modular structure of micro-kernels., executed at L0. Leveraging as L1 a GPH, like Xen or KVM, provides to the user an environment apparently  identical to the one used in production today, satisfying DP5. The L1 GPH is normally integrated with modern cloud toolkits like OpenStack and will support 3rd part applications.
Therefore, the architecture presents a stronger isolation leveraging on nested virtualization, providing in particular a strong tiny core(DP3). The MH layer will be transparent from the user and it will not not any change in the functioning of the platform.

@@

In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction. In this section, we propose a GPH over MH implementation, based on Xen over NOVA \cite{nova}.
